---
title: "COVID-19_virus_project"
author: "Kashish Gupta"
date: "2024-10-27"
output: html_document
---
              **Investigating COVID-19 Virus Trends**


#INTRODUCTION

This Project focuses on a pneumonia of unknown cause detected in Wuhan, China. It was first internationally reported from China on 31 December 2019. Today we know this virus as Coronavirus. COVID-19 which stands for COronaVIrus Disease is the disease caused by this virus. Since then, the world has been engaged in the fight against this pandemic. Several measures have therefore been taken to "flatten the curve". We have consequently experienced social distancing and many people have passed away as well.

In this Project, we use a dataset, from Kaggle. This dataset was collected between the 20th of January and the 1st of June 2020. 

#Purpose
The purpose of this Project is to provide an answer to this question:\ 
Which countries have reported the highest number of positive cases in relation to the number of tests conducted?

```{r load packages}
library(readr)
```

```{r import}
getwd()
setwd("C:/Users/HP/Downloads")

covid_df <- read_csv("covid19.csv")  

```

```{r}
dim(covid_df)
vector_cols <- colnames(covid_df)
View(vector_cols)

head(covid_df)
library(tibble)
glimpse(covid_df)
```

```{r load library}
library(dplyr)

#selecting only rows related to `All States` and removing the `Province_state`
covid_df_all_states <- covid_df %>%
  filter(Province_State == "All States") %>%
  select(-Province_State)


# Selecting the columns with cumulative numbers
covid_df_all_states_cumulative <- covid_df_all_states %>% 
  select(Date, Continent_Name, Two_Letter_Country_Code, positive, hospitalized, recovered, death, total_tested)

# Selecting the columns with cumulative numbers
covid_df_all_states_daily <- covid_df_all_states %>% 
  select(Date, Country_Region, active, hospitalizedCurr, daily_tested, daily_positive)

```

# Identifying the Highest Fatality Rates Countries
```{r}
## Summarizing the data based on the `Continent_Name` and `Two_Letter_Country_Code` columns.
covid_df_all_states_cumulative_max <- covid_df_all_states_cumulative %>% 
  group_by(Continent_Name, Two_Letter_Country_Code) %>% 
  summarise(max = max(death)) %>% 
  filter(max > 0)

covid_df_all_states_cumulative_max
```
## Displaying the maximum number of death by country, colored by continent
```{r}
library(ggplot2)

ggplot(data = covid_df_all_states_cumulative_max,aes(x = Two_Letter_Country_Code, y = max, col = Continent_Name)) + geom_point()
```
## Conclusion: Answering the question: Which countries have had the highest fatality (mortality) rates?
```{r}
death_top_3 <- c("US", "IT", "GB")
```

# Extracting the Top Ten countries in the number of tested cases
```{r}
covid_df_all_states_daily_sum <- covid_df_all_states_daily %>% 
  group_by(Country_Region)%>%
  summarize(tested = sum(daily_tested), positive = sum(daily_positive), active = sum(active), hospitalized = sum(hospitalizedCurr)) %>% arrange(desc(tested))

covid_df_all_states_daily_sum

covid_top_10 <- head(covid_df_all_states_daily_sum,10)
```
##Which countries have had the highest number of positive cases against the number of tests?
```{r}
# Identifying the Highest Positive Against Tested Cases
#vectors
countries <- covid_top_10$Country_Region 
tested_cases <- covid_top_10$tested
positive_cases <- covid_top_10$positive
active_cases <- covid_top_10$active
hospitalized_cases <- covid_top_10$hospitalized

#name the vectors
names(tested_cases) <- countries
names(positive_cases) <- countries
names(active_cases) <- countries
names(hospitalized_cases) <- countries

## Identifying 
positive_cases
sum(positive_cases)
mean(positive_cases)
positive_cases/sum(positive_cases)
```
```{r}
positive_cases/tested_cases

##conclusion
positive_tested_top_3 <- c("United Kingdom" = 0.11, "United States" = 0.10, "Turkey" = 0.08)
```
##Which countries have made the best effort in terms of the number of tests conducted related to their population?
# Identifying Affected Countries Related to their Population
```{r}
#Create a matrix combining the vectors:
covid_mat <- cbind(tested_cases, positive_cases, active_cases, hospitalized_cases)

population <- c(331002651, 145934462, 60461826, 1380004385, 84339067, 37742154, 67886011, 25499884, 32971854, 37846611)

covid_mat <- covid_mat * 100/population

covid_mat
```
# Ranking the matrix
```{r}
tested_cases_rank <- rank(covid_mat[,"tested_cases"])
positive_cases_rank <- rank(covid_mat[,"positive_cases"])
active_cases_rank <- rank(covid_mat[,"active_cases"])
hospitalized_cases_rank <- rank(covid_mat[,"hospitalized_cases"])

covid_mat_rank <- rbind(tested_cases_rank, positive_cases_rank, active_cases_rank, hospitalized_cases_rank)

covid_mat_rank

covid_mat_rank[1,]

covid_mat_rank[-1, ]

colSums(covid_mat_rank[-1, ])
```
##Which countries were ultimately the most and least affected related to their population?
```{r}
## Conclusion
best_effort_tested_cased_top_3 <- c("India", "United Kingdom", "Turkey")

most_affected_country <- "Italy"

least_affected_country <- "India"
```


```{r}
question_list <- list(
  "Which countries have had the highest fatality (mortality) rates?",
  "Which countries have had the highest number of positive cases against the number of tests?",
  "Which countries have made the best effort in terms of the number of tests conducted related to their population?",
  "Which countries were ultimately the most and least affected related to their population?"
)

answer_list <- list(
  "Death" = death_top_3,
  "Positive tested cases" = positive_tested_top_3,
  "The Best effort in test related to the population" = best_effort_tested_cased_top_3,
  "The most affected country related to its population" = most_affected_country,
  "The least affected country related to its population" = least_affected_country
)

answer_list

datasets <- list(
  original = covid_df,
  allstates = covid_df_all_states,
  cumulative = covid_df_all_states_cumulative,
  daily = covid_df_all_states_daily
)
matrices <- list(covid_mat, covid_mat_rank)
vectors <- list(vector_cols, population, countries)

data_structure_list <- list("data frame" = datasets, "matrix" = matrices, "vector" = vectors)

covid_analysis_list <- list(question_list, answer_list, data_structure_list)
```


